Dark matter halo shape in the Auriga simulations: radial dependence, time evolution and baryon disk alignment

### 2019-07-17 Document/paper.tex
edit

\title[The shape of dark matter haloes in the Auriga simulations]
{Dark matter halo shape in the Auriga simulations: radial
  dependence, time evolution and correlation with baryon disk
  properties}
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'sica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia.\\
$^{2}$Heidelberg Institute for Theoretical Studies,
Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany.\\
}

@@

\subsection{The effect of baryons at $z=0$}

@@

The situation in the MHD simulations is more diverse.
As expected by the results in the previous section the broad location
in the triaxility plane of the temporal and radial evolution in the
MHD simulation is completeley different from its DMO counterpart. 
In terms of the memory hypothesis we see two different trends.
We have the case of Halos \texttt{Au-6} and \texttt{Au-16}

\subsection{DM halo - Stellar Disk Alignment}

A common assumption in observational models of the MW DM halo is that
its minor axis is perfectly aligned with the disk axis.
Although it is a reasonable assumption to guarantee the
stability of the galactic disk in simplified models of isolated
galaxies, this might not hold in an explicit cosmological context.

To examine the validity of this assumption  we sampled the shape at 5
different radii and plotted  directions in a cartesian coordinate
system where the $z$-axis always corresponds to the minor axis
measured at the virial radius.

>>>>

\title[The shape of dark matter haloes in the Auriga simulations]
{Dark matter halo shape in the Auriga simulations: radial
  dependence, time evolution and baryon disk alignment}
\author[Jesus Prada,  Jaime E. Forero-Romero, Volker Springel ]{
Jesus Prada,$^{1}$\thanks{E-mail: jd.prada1760@uniandes.edu.co}
Jaime E. Forero-Romero,$^{1}$
Volker Springel$^{2}$
\\
% List of institutions
$^{1}$Departamento de F\'sica, Universidad de los Andes, Cra. 1 No.
18A-10, Edificio Ip, Bogot\'a, Colombia.\\
$^{2}$Heidelberg Institute for Theoretical Studies,
Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany.\\
}

@@

\subsection{Radial trends at $z=0$}

@@

The situation in the MHD simulations is more diverse.
As expected by the results in the previous section the broad location
in the triaxility plane of the temporal and radial evolution in the
MHD simulation is completeley different from its DMO counterpart. 
In terms of the correlations expected by the memory hypothesis we see
two different trends.
We have the case of Halos \texttt{Au-6}, \texttt{Au-16} and
\texttt{Au-27} there is still a reasonable correlation between
temporal and radial evolution. 
However for halos \texttt{Au-23} and \texttt{Au-24} the radial trends
go in directions that are not even covered by the temporal evolution.

Taken at face value, these results show that it is not possible to
generalize the memory hypotesis to dark matter halos in MHD
simulations.
This means that some halos present a amnesia to its formation history,
more specifically there are cases where the radial shape profile at
present time does not have any correlation to its evolutionary
history.

\subsection{Alignments with the stellar disk}

A common assumption in observational models of the MW DM halo is that
its minor axis is perfectly aligned with the stellar disk minor axis.
Although it is a reasonable assumption to guarantee the stability of
the galactic disk in simplified models of isolated galaxies, this
might not hold in an explicit cosmological context. 
To examine the degree of validity of this assumption we study in this
section the alignment between the eigenvectors of the inertia tensor of
stellar particles within $0.1R_{200}$ ($\approx 23\pm 2$ kpc) and the
eigenvectors of the dark matter halo shape.
